---
title: "Figure 4. Trophectoderm and placental-specific genes"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params:
  datadir: "./data/microscopy"
---

# Summary

This is the supplementary notebook for figure 4.


```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(rtracklayer)
library(tidyverse)
library(ggrastr)
library(DESeq2)
library(cowplot)
library(wigglescout)
library(readxl)
library(ggplot2)
library(ggpubr)
library(gtools)

source("./code/globals.R")
source("./code/embed_functions.R")
source("./code/heatmaply_functions.R")

knitr::opts_chunk$set(dev = c('png', 'my_svg'), fig.ext = c("png", "svg"), fig.width = 6, fig.height = 6)
genes <-
  read.table(
    "./data/meta/Kumar_2020_master_gene_table_rnaseq_shrunk_annotated.tsv",
    header = T,
    sep = "\t",
    colClasses = c(rep("character", 2), rep("numeric", 2), rep("factor", 3), rep("numeric", 86))
  )

dir.create("./figures_data")
dir.create("./output")

```

# Expression of trophectoderm and placental-specific genes

```{r lineage-markers-heatmap, message = F, warning = F}
genes_list <- c("EPAS1", "MSX2", "GATA3", "NR2F2", "CLDN4", "GATA2", "IGF2",
    "CDX2", "SLC40A1", "KRT7", "FRZB", "CGA", "ERP27", "KRT23", "CGB5", "VGLL1",
    "ENPEP", "TP63" )

fig <- combined_heatmap(genes, genes_list, cluster_rows = F,
  rnaseq_limits = c(0, 12.5),
  k4m3_limits = c(0, 80),
  k27m3_limits = c(0, 12),
  ub_limits = c(0, 12))

fig
```

# Image analysis

## 2 Inhibitors comparison

Channels:
-   c1: NANOG
-   c2: K27m3
-   c3: GATA3
-   c4: DAPI

`Metadata_Well`:

-   A05: Wildtype
-   B05: EZH2i_d7
-   C05: EEDi_d7

`Metadata_Site` represents a cluster of cells, where each cell has a number
assigned `ObjectNumber`.

```{r warning=F, message=F}
library(tidyverse)
library(ggplot2)
library(ggpubr)

source("./code/globals.R")

knitr::opts_chunk$set(error=FALSE, warning=FALSE, message=FALSE)
knitr::opts_chunk$set(dev = c('png', 'my_svg'), fig.ext = c("png", "svg"), fig.width = 8, fig.height = 8)
old <- theme_set(theme_classic())

expr <- read.table(file.path(params$datadir, "INH_well5_Nuclei.txt"), header = TRUE)

expr$Metadata_Well <-
  factor(
    expr$Metadata_Well,
    levels = c("A05", "B05", "C05"),
    labels = c("WT", "EZH2i", "EEDi")
  )

well <- filter(expr, expr$Intensity_MeanIntensity_c4_raw < 0.025 
               & Intensity_IntegratedIntensity_c4_raw < 30)
```

## H3K27m3 vs GATA3

```{r fig5c-channels-boxplot, fig.width = 7, fig.height = 3.5}
df <-
  well %>% dplyr::select(
    Metadata_Well,
    ObjectNumber,
    Metadata_Site,
    Intensity_MeanIntensity_c1_raw,
    Intensity_MeanIntensity_c2_raw,
    Intensity_MeanIntensity_c3_raw,
    Intensity_MeanIntensity_c4_raw
  ) %>%
  rename(
    NANOG = Intensity_MeanIntensity_c1_raw,
    H3K27m3 = Intensity_MeanIntensity_c2_raw,
    GATA3 = Intensity_MeanIntensity_c3_raw,
    DAPI = Intensity_MeanIntensity_c4_raw,
    Well = Metadata_Well,
    Cell = ObjectNumber,
    Site = Metadata_Site
  )

df_long <- df %>% pivot_longer(!c(Well, Cell, Site), names_to = "channel", values_to = "value")

my_comparisons <- list(c("EZH2i", "WT"), c("EEDi", "WT"))

ggplot(df_long %>% filter(channel != "DAPI"), aes(y=value, x=Well, color=Well)) + 
  geom_boxplot() + 
  facet_wrap(. ~ channel, scales = "free_y", nrow = 1) + 
  stat_compare_means(comparisons = my_comparisons, method = "t.test") +
  scale_color_manual(values = c("grey", "#4682B4", "#8B0000"))


summaries <- df %>% group_by(Well) %>% summarise(cutoff = mean(GATA3)*1.5)
cutoff <- summaries[summaries$Well == "WT", "cutoff"][[1]]
```


## GATA3 vs NANOG and H3K27m3

GATA3 vs H3K27m3 per nucleus

```{r fig5d-gata3-vs-h3k27m3, fig.width = 6, fig.height = 6}
ggplot(df, aes(x = H3K27m3, y = GATA3, color = Well)) + 
  geom_point(alpha = 0.7, size = 2) + 
  scale_color_manual(values = c("grey", "#4682B4", "#8B0000"))  + 
  geom_hline(yintercept =cutoff, linetype = "dotted")

```

GATA3 vs NANOG per nucleus

```{r fig5d-gata3-vs-nanog, fig.width = 6, fig.height = 6}
ggplot(df, aes(x = NANOG, y = GATA3, color = Well)) + 
  geom_point(alpha = 0.7, size = 2) + 
  scale_color_manual(values = c("grey", "#4682B4", "#8B0000")) + 
  geom_hline(yintercept =cutoff, linetype = "dotted")

```

```{r calculate-perc-cells}
# Total number of cells
perc_positive <- right_join(df %>% group_by(Well) %>% summarise(n_total = n()),
                            df %>% group_by(Well) %>% filter(GATA3 > cutoff) %>% 
                              summarise(n_positive = n()), by="Well") %>%
  mutate(perc_positive = (n_positive / n_total)*100)

perc_positive
```


## H3K27m3 Timecourse

Channels:

-   c1: NANOG
-   c2: H3K37m3
-   c3: OCT3/4
-   c4: DAPI

`Metadata_Well`:

-   A05: Wildtype
-   B05: EZH2i_d2
-   C05: EZH2i_d4
-   D05: EZH2i_d7

```{r warning=F, message=F}
expr <- read.table(file.path(params$datadir, "./Timecourse_Well5Nuclei.txt") ,header = TRUE)

expr$Metadata_Well <- factor(
    expr$Metadata_Well,
    levels = c("A05", "B05", "C05", "D05"),
    labels = c("WT", "EZH2i_D2", "EZH2i_D4", "EZH2i_D7")
  )

df <-
  expr %>% dplyr::select(
    Metadata_Well,
    ObjectNumber,
    Metadata_Site,
    Intensity_MeanIntensity_c1_raw,
    Intensity_MeanIntensity_c2_raw,
    Intensity_MeanIntensity_c3_raw,
    Intensity_MeanIntensity_c4_raw
  ) %>%
  rename(
    NANOG = Intensity_MeanIntensity_c1_raw,
    H3K27m3 = Intensity_MeanIntensity_c2_raw,
    OCT = Intensity_MeanIntensity_c3_raw,
    DAPI = Intensity_MeanIntensity_c4_raw,
    Well = Metadata_Well,
    Cell = ObjectNumber,
    Site = Metadata_Site
  )
```


```{r fig5e-h3k27m3-timecourse, fig.width = 5, fig.height = 5}
df_long <- df %>% pivot_longer(c(NANOG, H3K27m3, OCT, DAPI),
                               names_to = "channel", values_to = "value")

my_comparisons <- list(c("EZH2i_D2", "WT"), c("EZH2i_D4", "WT"), c("EZH2i_D7", "WT"))

ggplot(df_long %>% filter(channel == "H3K27m3"),
       aes(y=value, x=Well, color = channel)) + 
  geom_boxplot() + 
  stat_compare_means(comparisons = my_comparisons, method = "t.test") +
  labs(title = "Time course - mean H3K27m3 intensity", y = "Mean intensity", x = "") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_color_manual(values = c("navy"))
```

## CRISPR/Cas targeting 

```{r merge-crispr-rounds}
g1_eed <- read.csv(file.path(params$datadir, "g1_eed.csv"))
g2_eed <- read.csv(file.path(params$datadir, "g2_eed.csv"))

median_1 <- median(g1_eed[g1_eed$Well == "WT", "EED"])
median_2 <- median(g2_eed[g2_eed$Well == "NTgRNA", "EED"])

median_1_gata3 <- median(g1_eed[g1_eed$Well == "WT", "GATA3"])
median_2_gata3 <- median(g2_eed[g2_eed$Well == "NTgRNA", "GATA3"])

merged_rounds <- 
  rbind(g1_eed %>% mutate(round = "1", EED_norm = EED / median_1, GATA3_norm = GATA3 / median_1_gata3), 
        g2_eed %>% mutate(round = "2", EED_norm = EED / median_2, GATA3_norm = GATA3 / median_2_gata3) %>% 
  dplyr::select(!contains("GATA6")))
```


```{r calculate-positive-crispr}
library(ggrastr)
merged_rounds <- merged_rounds %>% mutate(condition = ifelse(Well == "Transfected", "Transfected", "Control"))

summaries <- merged_rounds %>% 
  group_by(condition) %>% summarise(cutoff = mean(GATA3_norm)*1.5)

cutoff <- summaries[summaries$condition == "Control", "cutoff"][[1]]

# Total number of cells
perc_positive <- right_join(
  merged_rounds %>% group_by(condition, round) %>% summarise(n_total = n()),
  merged_rounds %>% group_by(condition, round) %>% filter(GATA3_norm > cutoff) %>%
    summarise(n_positive = n())) %>%
  mutate(perc_positive = (n_positive / n_total)*100)

perc_positive
```

```{r fig5g-crispr-cas-targeting, fig.height = 5, fig.width = 5}
ggplot(merged_rounds, 
       aes(x = EED_norm, y = GATA3_norm, color = interaction(Well, round))) + 
  rasterise(geom_point(size = 0.7, alpha = 0.5), dpi = 300) + 
  scale_color_manual(values = c("#f58e8e", "grey", "#99AABB", "#CD5C5C")) +
  labs(title = "Normalized mean intensity across rounds") +
  theme(legend.title = element_blank()) + geom_hline(yintercept = cutoff, linetype = "dotted")
```


```{r boxplot-merged-crispr-facet}
mr_long <- merged_rounds %>% select(Well, Site, round, EED_norm, GATA3_norm) %>% pivot_longer(c(EED_norm, GATA3_norm), names_to = "channel", values_to = "norm_intensity") %>% mutate(condition = ifelse(Well == "Transfected", "Transfected", "Control"))

my_comp <- list(c("Transfected", "Control"))
mr_long$Site <- as.factor(mr_long$Site)
ggplot(mr_long %>% filter(channel == "GATA3_norm"), aes(x=Site, y=norm_intensity, color = condition)) + 
  geom_boxplot() + facet_wrap(round ~ condition)
```


## Extended 10a. HS975

```{r warning=F, message=F}
old <- theme_set(theme_classic())

expr <- read.table(file.path(params$datadir, "HS975_20xNuclei.txt"), header = TRUE)
expr$Metadata_Site <- sapply(expr$Metadata_Site, as.factor)

# Rename wells for readability
expr$Metadata_Well <- factor(expr$Metadata_Well, levels = c("A01", "B01"),
                             labels = c("WT", "EZH2i"))
well <- filter(expr, expr$Intensity_MeanIntensity_c4_raw >= 0.025)
```


```{r}
df <-
  well %>% select(
    Metadata_Well,
    ObjectNumber,
    Metadata_Site,
    Intensity_MeanIntensity_c1_raw,
    Intensity_MeanIntensity_c2_raw,
    Intensity_MeanIntensity_c3_raw,
    Intensity_MeanIntensity_c4_raw
  ) %>%
  rename(
    GATA6 = Intensity_MeanIntensity_c1_raw,
    H3K27m3 = Intensity_MeanIntensity_c2_raw,
    GATA3 = Intensity_MeanIntensity_c3_raw,
    DAPI = Intensity_MeanIntensity_c4_raw,
    Well = Metadata_Well,
    Cell = ObjectNumber,
    Site = Metadata_Site
  )

noise_cutoff <- 0.007
df <- df %>% filter(GATA3 > noise_cutoff)
```

```{r perc-positive-hs975}
summaries <- df %>% 
  group_by(Well) %>% summarise(cutoff = mean(GATA3)*1.5)
cutoff <- summaries[summaries$Well == "WT", "cutoff"][[1]]
perc_positive <- right_join(df %>% group_by(Well) %>% summarise(n_total = n()),
                            df %>% group_by(Well) %>% filter(GATA3 > cutoff) %>%
                              summarise(n_positive = n())) %>%
  mutate(perc_positive = (n_positive / n_total)*100)

perc_positive
```

```{r extfig10-boxplot}
df_long <- df %>% pivot_longer(!c(Well, Cell, Site),
                               names_to = "channel", values_to = "value")

ggplot(df_long %>% filter(channel %in% c("GATA3", "H3K27m3")), aes(y=value, x=Well, color=Well)) + 
  geom_boxplot() + facet_grid(. ~ channel, scales = "free_y") + 
  stat_compare_means(method="t.test")
```

