---
title: "Figure 2. H3K27m3 Naïve and Primed-specific bivalency groups"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params:
  output_matrices: FALSE
---

# Summary

This is the supplementary notebook for figure 3.

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(rtracklayer)
library(tidyverse)
library(effsize)
library(ggpubr)
library(ggplot2)
library(ggrastr)
library(cowplot)
library(ggalluvial)
library(dplyr)

source("./code/globals.R")
source("./code/embed_functions.R")
source("./code/heatmap_panels.R")
source("./code/heatmaply_functions.R")

knitr::opts_chunk$set(dev = c('png', 'my_svg'), fig.ext = c("png", "svg"), fig.width = 5, fig.height = 5)
knitr::opts_chunk$set(class.source='fold-show')

genes <-
  read.table(
    "./data/meta/Kumar_2020_master_gene_table_rnaseq_shrunk_annotated.tsv",
    header = T,
    sep = "\t",
    colClasses = c(rep("character", 5), rep("factor", 3), rep("numeric", 86))
  )

make_gr_from_table <- function(df) {
  makeGRangesFromDataFrame(df[, c("seqnames", "start", "end", "strand", "name")], keep.extra.columns = T)
}

dir.create("./figures_data")
dir.create("./output")

pval_cutoff <- 0.05
fc_cutoff <- log2(1.5)

```

# H3K27m3 groups 

## H3K27m3 group selection

```{r group-selection}
k27_groups <- split(genes, genes$k27_bivalency_grp)
k27_groups_loci <- lapply(k27_groups, make_gr_from_table)
```

## H3K27m3 alluvial plot

```{r alluvial-plot-histone-conditions, warning = F, message = F}

gr_annot <-
  genes %>% dplyr::select(name, seqnames, start, end, strand,
    H3K4m3_DS_Pr_vs_Ni_log2FoldChange,
    H3K27m3_DS_Pr_vs_Ni_log2FoldChange,
    H2Aub_DS_Pr_vs_Ni_log2FoldChange,
    H3K4m3_DS_Pr_vs_Ni_padj,
    H3K27m3_DS_Pr_vs_Ni_padj,
    H2Aub_DS_Pr_vs_Ni_padj,
    k27_bivalency_grp
  )

gr_annot <- gr_annot %>% filter(k27_bivalency_grp %in% c("Always_up", "Ni_higher_than_Pr", "Pr_higher_than_Ni")) %>% mutate(
   k4 = factor(case_when(
     H3K4m3_DS_Pr_vs_Ni_padj < pval_cutoff & H3K4m3_DS_Pr_vs_Ni_log2FoldChange > fc_cutoff ~ "Pr >> Ni",
     H3K4m3_DS_Pr_vs_Ni_padj < pval_cutoff & H3K4m3_DS_Pr_vs_Ni_log2FoldChange < -fc_cutoff ~ "Ni >> Pr",
     TRUE ~ "unchanged")
   ),
   k27 = factor(case_when(
     H3K27m3_DS_Pr_vs_Ni_padj < pval_cutoff & H3K27m3_DS_Pr_vs_Ni_log2FoldChange > fc_cutoff ~ "Pr >> Ni",
     H3K27m3_DS_Pr_vs_Ni_padj < pval_cutoff & H3K27m3_DS_Pr_vs_Ni_log2FoldChange < -fc_cutoff ~ "Ni >> Pr",
     TRUE ~ "unchanged")
   ),
   ub = factor(case_when(
     H2Aub_DS_Pr_vs_Ni_padj < pval_cutoff & H2Aub_DS_Pr_vs_Ni_log2FoldChange > fc_cutoff ~ "Pr >> Ni",
     H2Aub_DS_Pr_vs_Ni_padj < pval_cutoff & H2Aub_DS_Pr_vs_Ni_log2FoldChange < -fc_cutoff ~ "Ni >> Pr",
     TRUE ~ "unchanged")
   ))

counts_table <- gr_annot %>% dplyr::count(k4, k27)

ggplot(counts_table, aes(y = n, axis1 = k4, axis2 = k27, fill = k27)) +
  geom_alluvium(width = 1 / 12) +
  geom_stratum(width = 1 / 12, color = "black") + 
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), nudge_x = -0.08, angle = 90, size = 3) + 
  scale_fill_manual(values = c("#278b8b",  "#f44b34", "#cccccc")) +
  theme_default() + 
  scale_x_discrete(limits = c("H3K4m3", "H3K27m3"), expand = c(.05, .05)) 
```

# H3K27m3 group heatmap panels

## H3K27m3 and H2AUb

```{r heatmap-groups-1, warning=F, fig.width = 14, fig.height = 14}
plot_bw_heatmap_panel(
    c(bwfiles$k27[c(1, 3)], bwfiles$ub[c(1, 3)]),
    list(k27_groups_loci$Pr_higher_than_Ni,
         k27_groups_loci$Ni_higher_than_Pr,
         k27_groups_loci$Always_up
    ),
    c("H3K27m3_Ni", "H3K27m3_Pr", "H2Aub_Ni", "H2Aub_Pr"),
    c("Primed >> Naive", "Naive >> Primed", "Always up"),
    global_scale = TRUE,
    proportional = TRUE,
    mode = "center"
)
```

Plot the second part of the heatmap panel.

```{r heatmap-groups-2, warning=F, fig.width = 14, fig.height = 10}
plot_bw_heatmap_panel(
  c(bwfiles$k27[c(1, 3)], bwfiles$ub[c(1, 3)]),
  list(k27_groups_loci$Pr_higher_than_Ni, k27_groups_loci$K4_only),
  c("H3K27m3_Ni", "H3K27m3_Pr", "H2Aub_Ni", "H2Aub_Pr"),
  c("Primed >> Naive", "Rest"),
  global_scale = TRUE,
  proportional = TRUE,
  mode = "center",
  zmin = 0,
  zmax = 23
)

```


## H3K4m3

Plot the H3K4m3 part, sorted by the same reference

```{r heatmap-groups-3, warning=F, fig.width = 14, fig.height = 14}
plot_bw_heatmap_panel(
    c(bwfiles$k27[1], bwfiles$k4[c(1, 3)]),
    list(k27_groups_loci$Pr_higher_than_Ni,
         k27_groups_loci$Ni_higher_than_Pr,
         k27_groups_loci$Always_up
    ),
    c("H3K27m3_Ni", "H3K4m3_Ni", "H3K4m3_Pr"),
    c("Primed >> Naive", "Naive >> Primed", "Always up"),
    global_scale = TRUE,
    proportional = TRUE,
    mode = "center",
    zmin = 0,
    zmax = 188
)
```


```{r heatmap-groups-4, warning=F, fig.width = 14, fig.height = 10}
plot_bw_heatmap_panel(
  c(bwfiles$k27[1], bwfiles$k4[c(1, 3)]),
  list(k27_groups_loci$Pr_higher_than_Ni, k27_groups_loci$K4_only),
  c("H3K27m3_Ni", "H3K4m3_Ni", "H2Aub_Pr"),
  c("Primed >> Naive", "Rest"),
  global_scale = TRUE,
  proportional = TRUE,
  mode = "center"
)

```


```{r heatmap-panel-data-tables, warning = F, message = F, class.source = NULL, echo = F}
# This auxiliary code creates files with the actual values in the profiles and
# heatmap matrices. It is not output in the report for simplicity, but kept
# here for completeness.

if (params$output_matrices == TRUE) {
    
  write_mat <- function(m, f) {
    write.table(m, file = f, quote = F, sep = "\t", row.names = F)
  }
  
  select_order <- function(m, order) {
    m[[1]][order, ]
  }
  
  bw_used <- c(bwfiles$k27[c(1,3)], bwfiles$k4[c(1,3)], bwfiles$ub[c(1,3)])
  bw_ref <- bwfiles$k27[1]
  
  heatmap_func <- purrr::partial(bw_heatmap, mode = "center", upstream = 2500, downstream = 2500)
  
  # Primed >> Naive
  heatmap_ref <- heatmap_func(bw_ref, loci = k27_groups_loci$up)
  order_ref <- order(rowMeans(heatmap_ref[[1]]), decreasing = T)
  
  mat_list <- lapply(bw_used, heatmap_func, loci = k27_groups_loci$up)
  # Reorder according to reference
  mat_list_sorted <- lapply(mat_list, select_order, order = order_ref)
  
  result <- purrr::map2(mat_list_sorted, 
              file.path("./figures_data",
              paste0("fig2_heatmap_", gsub(".hg38.scaled.bw", "", basename(bw_used)), "_Pr_higher_Ni.tsv")),
              write_mat)
  
  # Naive >> Primed
  heatmap_ref <- heatmap_func(bw_ref, loci = k27_groups_loci$Ni_higher_than_Pr)
  order_ref <- order(rowMeans(heatmap_ref[[1]]), decreasing = T)
  
  mat_list <- lapply(bw_used, heatmap_func, loci = k27_groups_loci$Ni_higher_than_Pr)
  # Reorder according to reference
  mat_list_sorted <- lapply(mat_list, select_order, order = order_ref)
  
  result <- purrr::map2(mat_list_sorted, 
              file.path("./figures_data", paste0("fig2_heatmap_", gsub(".hg38.scaled.bw", "", basename(bw_used)), "_Ni_higher_Pr.tsv")),
              write_mat)
  
  
  # Always up
  heatmap_ref <- heatmap_func(bw_ref, loci = k27_groups_loci$Always_up)
  order_ref <- order(rowMeans(heatmap_ref[[1]]), decreasing = T)
  
  mat_list <- lapply(bw_used, heatmap_func, loci = k27_groups_loci$Always_up)
  # Reorder according to reference
  mat_list_sorted <- lapply(mat_list, select_order, order = order_ref)
  
  result <- purrr::map2(mat_list_sorted, 
              file.path("./figures_data", paste0("fig2_heatmap_", gsub(".hg38.scaled.bw", "", basename(bw_used)), "_always_up.tsv")),
              write_mat)
  
  # Rest
  heatmap_ref <- heatmap_func(bw_ref, loci = k27_groups_loci$K4_only)
  order_ref <- order(rowMeans(heatmap_ref[[1]]), decreasing = T)
  
  mat_list <- lapply(bw_used, heatmap_func, loci = k27_groups_loci$K4_only)
  # Reorder according to reference
  mat_list_sorted <- lapply(mat_list, select_order, order = order_ref)
  
  result <- purrr::map2(mat_list_sorted, 
              file.path("./figures_data", paste0("fig2_heatmap_", gsub(".hg38.scaled.bw", "", basename(bw_used)), "_not_enriched.tsv")),
              write_mat)
  
  # Profiles
  write_mat(
    bw_profile(
      bw_used,
      loci = k27_groups_loci$Ni_higher_than_Pr,
      mode = "center",
      upstream = 2500,
      downstream = 2500
    ),
    "./figures_data/fig2_profile_all_Ni_higher_Pr.tsv"
  )
  write_mat(
    bw_profile(
      bw_used,
      loci = k27_groups_loci$Pr_higher_than_Ni,
      mode = "center",
      upstream = 2500,
      downstream = 2500
    ),
    "./figures_data/fig2_profile_all_Pr_higher_Ni.tsv"
  )
  write_mat(
    bw_profile(
      bw_used,
      loci = k27_groups_loci$Always_up,
      mode = "center",
      upstream = 2500,
      downstream = 2500
    ),
    "./figures_data/fig2_profile_all_always_up.tsv"
  )
  write_mat(
    bw_profile(
      bw_used,
      loci = k27_groups_loci$K4_only,
      mode = "center",
      upstream = 2500,
      downstream = 2500
    ),
    "./figures_data/fig2_profile_all_not_enriched.tsv"
  )
}


```


# RNA-seq ratios at H3K27m3 groups

## H327m3 Naive >> Primed

```{r rnaseq-ratio-violin-h3k27m3-ni-higher, warning = F, message = F}

df <-
  genes[, c(
    "name",
    "RNASeq_DS_Pr_vs_Ni_log2FoldChange",
    "RNASeq_DS_EZH2i_vs_Ni_log2FoldChange",
    "RNASeq_DS_EZH2i_vs_Pr_log2FoldChange"
  )]

colnames(df) <- c("name", "Primed vs Naive", "EZH2i vs Naive", "EZH2i vs Primed")
df_k27_up <- df[df$name %in% k27_groups_loci$Ni_higher_than_Pr$name, ]

df_long <- df %>% pivot_longer(!name, names_to = "group", values_to = "fc")
df_long_k27 <- df_k27_up %>% pivot_longer(!name, names_to = "group", values_to = "fc")

df_long$group <- factor(df_long$group, levels = c("Primed vs Naive", "EZH2i vs Naive", "EZH2i vs Primed"))
df_long_k27$group <- factor(df_long_k27$group, levels = c("Primed vs Naive", "EZH2i vs Naive", "EZH2i vs Primed"))

# Stats subset vs global
ni_pr_test <- wilcox.test(df_k27_up[, "Primed vs Naive"],
  df[!df$name %in% df_k27_up$name , "Primed vs Naive"], na.rm = T)

ni_pr_effect <- cohen.d(df_k27_up[, "Primed vs Naive"],
  df[!df$name %in% df_k27_up$name , "Primed vs Naive"], na.rm = T)

col <- "EZH2i vs Naive"
ni_ezh2i_test <- wilcox.test(df_k27_up[, col],
  df[!df$name %in% df_k27_up$name , col], na.rm = T)

ni_ezh2i_effect <- cohen.d(df_k27_up[, col],
  df[!df$name %in% df_k27_up$name , col], na.rm = T)

col <- "EZH2i vs Primed"
pr_ezh2i_test <- wilcox.test(df_k27_up[, col],
  df[!df$name %in% df_k27_up$name , col], na.rm = T)

pr_ezh2i_effect <- cohen.d(df_k27_up[, col],
  df[!df$name %in% df_k27_up$name , col], na.rm = T)

stats_caption <- paste("Wilcoxon global vs. selection",
  paste("Ni_Pr:", format(ni_pr_test$p.value, digits=8),
        "Cohen D:", round(ni_pr_effect$estimate, digits = 4),
        "(", ni_pr_effect$magnitude, ")"),
  paste("Ni_EZH2i:", format(ni_ezh2i_test$p.value, digits=8),
        "Cohen D:", round(ni_ezh2i_effect$estimate, digits = 4),
        "(", ni_ezh2i_effect$magnitude, ")"),
  paste("Pr_EZH2i:", format(pr_ezh2i_test$p.value, digits=8),
        "Cohen D:", round(pr_ezh2i_effect$estimate, digits = 4),
        "(", pr_ezh2i_effect$magnitude, ")"),
  sep = "\n")

my_comparisons <- list(c("EZH2i vs Naive", "EZH2i vs Primed"),
                       c("Primed vs Naive", "EZH2i vs Naive"),
                       c("Primed vs Naive", "EZH2i vs Primed"))

k27_up_color <- "#009784"

ggplot(data=df_long, aes(x = group, y = fc)) +
  geom_violin(size = 0.8) +
  stat_compare_means(data=df_long_k27,
    comparisons = my_comparisons, method = "wilcox.test", paired = TRUE) +
  rasterize(
    geom_jitter(data=df_long_k27, color = k27_up_color, alpha = 0.5, size = 0.1),
    dpi = 300) +
  geom_hline (yintercept = 0, linetype = "dashed") + 
  coord_cartesian(ylim = c(-13, 24)) +
  theme_default(base_size = 14) + 
  labs(x = "Condition",
       y = "Log2 FC",
       title = "RNASeq expression changes",
       subtitle = "Global vs H3K27m3 Naïve >> Primed TSS",
       caption = stats_caption)

```

Download values: `r embed_df(df_long[!is.na(df_long$fc), ], name="log2fc_distribution.tsv", text = "Distribution")`, 
`r embed_df(df_long_k27[!is.na(df_long_k27$fc), ], name="k27_ni_higher_jitter.tsv", text="K27 higher naive points")`, 

## H327m3 Primed >> Naive

```{r rnaseq-ratio-violin-h3k27m3-pr-higher, warning = F, message = F}

df <-
  genes[, c(
    "name",
    "RNASeq_DS_Pr_vs_Ni_log2FoldChange",
    "RNASeq_DS_EZH2i_vs_Ni_log2FoldChange",
    "RNASeq_DS_EZH2i_vs_Pr_log2FoldChange"
  )]

colnames(df) <- c("name", "Primed vs Naive", "EZH2i vs Naive", "EZH2i vs Primed")

df_k27_up <- df[df$name %in% k27_groups_loci$Pr_higher_than_Ni$name, ]

df_k27_up$`Primed vs Naive` <- -df_k27_up$`Primed vs Naive`
df$`Primed vs Naive` <- -df$`Primed vs Naive`

df_long <- df %>% pivot_longer(!name, names_to = "group", values_to = "fc")
df_long_k27 <- df_k27_up %>% pivot_longer(!name, names_to = "group", values_to = "fc")

df_long$group <- factor(df_long$group, levels = c("Primed vs Naive", "EZH2i vs Naive", "EZH2i vs Primed"))
df_long_k27$group <- factor(df_long_k27$group, levels = c("Primed vs Naive", "EZH2i vs Naive", "EZH2i vs Primed"))

levels(df_long$group) <- c("Naive vs Primed", "EZH2i vs Naive", "EZH2i vs Primed")
levels(df_long_k27$group) <- c("Naive vs Primed", "EZH2i vs Naive", "EZH2i vs Primed")

my_comparisons <- list(c("EZH2i vs Naive", "EZH2i vs Primed"),
                       c("Naive vs Primed", "EZH2i vs Naive"),
                       c("Naive vs Primed", "EZH2i vs Primed"))

# Stats subset vs global
ni_pr_test <- wilcox.test(df_k27_up[, "Primed vs Naive"],
  df[!df$name %in% df_k27_up$name , "Primed vs Naive"], na.rm = T)

ni_pr_effect <- cohen.d(df_k27_up[, "Primed vs Naive"],
  df[!df$name %in% df_k27_up$name , "Primed vs Naive"], na.rm = T)

col <- "EZH2i vs Naive"
ni_ezh2i_test <- wilcox.test(df_k27_up[, col],
  df[!df$name %in% df_k27_up$name , col], na.rm = T)

ni_ezh2i_effect <- cohen.d(df_k27_up[, col],
  df[!df$name %in% df_k27_up$name , col], na.rm = T)

col <- "EZH2i vs Primed"
pr_ezh2i_test <- wilcox.test(df_k27_up[, col],
  df[!df$name %in% df_k27_up$name , col], na.rm = T)

pr_ezh2i_effect <- cohen.d(df_k27_up[, col],
  df[!df$name %in% df_k27_up$name , col], na.rm = T)

stats_caption <- paste("Wilcoxon global vs. selection",
  paste("Ni_Pr:", format(ni_pr_test$p.value, digits=8),
        "Cohen D:", round(ni_pr_effect$estimate, digits = 4),
        "(", ni_pr_effect$magnitude, ")"),
  paste("Ni_EZH2i:", format(ni_ezh2i_test$p.value, digits=8),
        "Cohen D:", round(ni_ezh2i_effect$estimate, digits = 4),
        "(", ni_ezh2i_effect$magnitude, ")"),
  paste("Pr_EZH2i:", format(pr_ezh2i_test$p.value, digits=8),
        "Cohen D:", round(pr_ezh2i_effect$estimate, digits = 4),
        "(", pr_ezh2i_effect$magnitude, ")"),
  sep = "\n")

k27_up_color <- "#ff4b40"

ggplot(data=df_long, aes(x = group, y = fc)) +
  geom_violin(size = 0.8) +
  stat_compare_means(
    data=df_long_k27, comparisons = my_comparisons,
    method = "wilcox.test", paired = TRUE) +
  rasterize(
    geom_jitter(data=df_long_k27, color = k27_up_color, alpha = 0.5, size = 0.1),
    dpi = 300) +
  geom_hline (yintercept = 0, linetype = "dashed") + 
  coord_cartesian(ylim = c(-13, 24)) +
  theme_default(base_size = 14) + 
  labs(x = "Condition",
       y = "Log2 FC",
       title = "RNASeq expression changes",
       subtitle = "Global vs H3K27m3 Primed >> Naïve TSS",
       caption = stats_caption)


write.table(df_long[!is.na(df_long$fc), ],
  file = "./figures_data/fig2_violin_k27_pr_higher_violin.tsv",
  col.names = T, sep = "\t", quote = F, row.names = F)

write.table(df_long_k27[!is.na(df_long_k27$fc), ],
  file = "./figures_data/fig2_violin_k27_pr_higher_jitter_k27_values.tsv",
  col.names = T, sep = "\t", quote = F, row.names = F)


```

Download values: `r embed_df(df_long_k27[!is.na(df_long_k27$fc), ], name="k27_pr_higher_jitter.tsv", text = "K27 primed higher points")`. 



## H327m3 always high

```{r rnaseq-ratio-violin-h3k27m3-always-up, warning = F, message = F}

df <-
  genes[, c(
    "name",
    "RNASeq_DS_Pr_vs_Ni_log2FoldChange",
    "RNASeq_DS_EZH2i_vs_Ni_log2FoldChange",
    "RNASeq_DS_EZH2i_vs_Pr_log2FoldChange"
  )]

colnames(df) <- c("name", "Primed vs Naive", "EZH2i vs Naive", "EZH2i vs Primed")

df_k27_up <- df[df$name %in% k27_groups_loci$Always_up$name, ]

df_k27_up$`Primed vs Naive` <- -df_k27_up$`Primed vs Naive`
df$`Primed vs Naive` <- -df$`Primed vs Naive`

df_long <- df %>% pivot_longer(!name, names_to = "group", values_to = "fc")
df_long_k27 <- df_k27_up %>% pivot_longer(!name, names_to = "group", values_to = "fc")

df_long$group <- factor(df_long$group, levels = c("Primed vs Naive", "EZH2i vs Naive", "EZH2i vs Primed"))
df_long_k27$group <- factor(df_long_k27$group, levels = c("Primed vs Naive", "EZH2i vs Naive", "EZH2i vs Primed"))

levels(df_long$group) <- c("Naive vs Primed", "EZH2i vs Naive", "EZH2i vs Primed")
levels(df_long_k27$group) <- c("Naive vs Primed", "EZH2i vs Naive", "EZH2i vs Primed")

my_comparisons <- list(c("EZH2i vs Naive", "EZH2i vs Primed"),
                       c("Naive vs Primed", "EZH2i vs Naive"),
                       c("Naive vs Primed", "EZH2i vs Primed"))

# Stats subset vs global
ni_pr_test <- wilcox.test(df_k27_up[, "Primed vs Naive"],
  df[!df$name %in% df_k27_up$name , "Primed vs Naive"], na.rm = T)

ni_pr_effect <- cohen.d(df_k27_up[, "Primed vs Naive"],
  df[!df$name %in% df_k27_up$name , "Primed vs Naive"], na.rm = T)

col <- "EZH2i vs Naive"
ni_ezh2i_test <- wilcox.test(df_k27_up[, col],
  df[!df$name %in% df_k27_up$name , col], na.rm = T)

ni_ezh2i_effect <- cohen.d(df_k27_up[, col],
  df[!df$name %in% df_k27_up$name , col], na.rm = T)

col <- "EZH2i vs Primed"
pr_ezh2i_test <- wilcox.test(df_k27_up[, col],
  df[!df$name %in% df_k27_up$name , col], na.rm = T)

pr_ezh2i_effect <- cohen.d(df_k27_up[, col],
  df[!df$name %in% df_k27_up$name , col], na.rm = T)

stats_caption <- paste("Wilcoxon global vs. selection",
  paste("Ni_Pr:", format(ni_pr_test$p.value, digits=8),
        "Cohen D:", round(ni_pr_effect$estimate, digits = 4),
        "(", ni_pr_effect$magnitude, ")"),
  paste("Ni_EZH2i:", format(ni_ezh2i_test$p.value, digits=8),
        "Cohen D:", round(ni_ezh2i_effect$estimate, digits = 4),
        "(", ni_ezh2i_effect$magnitude, ")"),
  paste("Pr_EZH2i:", format(pr_ezh2i_test$p.value, digits=8),
        "Cohen D:", round(pr_ezh2i_effect$estimate, digits = 4),
        "(", pr_ezh2i_effect$magnitude, ")"),
  sep = "\n")

k27_up_color <- "#4463b9"

ggplot(data=df_long, aes(x = group, y = fc)) +
  geom_violin(size = 0.8) +
  stat_compare_means(
    data=df_long_k27, comparisons = my_comparisons,
    method = "wilcox.test", paired = TRUE) +
  rasterize(
    geom_jitter(data=df_long_k27, color = k27_up_color, alpha = 0.5, size = 0.1),
    dpi = 300) +
  geom_hline (yintercept = 0, linetype = "dashed") + 
  coord_cartesian(ylim = c(-13, 24)) +
  theme_default(base_size = 14) + 
  labs(x = "Condition",
       y = "Log2 FC",
       title = "RNASeq expression changes",
       subtitle = "Global vs H3K27m3 H3K27m3 always up TSS",
       caption = stats_caption)


write.table(df_long[!is.na(df_long$fc), ],
  file = "./figures_data/fig2_violin_k27_always_up_violin.tsv",
  col.names = T, sep = "\t", quote = F, row.names = F)

write.table(df_long_k27[!is.na(df_long_k27$fc), ],
  file = "./figures_data/fig2_violin_k27_always_up_jitter_k27_values.tsv",
  col.names = T, sep = "\t", quote = F, row.names = F)


```

Download values: `r embed_df(df_long_k27[!is.na(df_long_k27$fc), ], name="k27_always_up_jitter.tsv", text = "K27 always up points")`. 


# Combined heatmaps

## Naive markers

```{r combined-summarized-heatmaps-naive, warning=F, message=F, fig.width = 7, fig.height = 8}
naive_markers <-
  c("KLF17",
    "DPPA5",
    "DNMT3L",
    "GATA6",
    "TBX3",
    "IL6ST",
    "DPPA3",
    "KLF5",
    "KLF4",
    "HORMAD1",
    "KHDC3L",
    "ALPP",
    "ALPPL2",
    "ZNF729",
    "TRIM60",
    "TFAP2C"
  )

tpm_cols <- grep("RNASeq_TPM", colnames(genes), value = T)
cols_used <- c("name", "H3K27m3_Ni_mean_cov", "H3K27m3_Pr_mean_cov", "H3K27m3_Ni_EZH2i_mean_cov", "H3K27m3_Pr_EZH2i_mean_cov",
               "H3K4m3_Ni_mean_cov", "H3K4m3_Pr_mean_cov", "H3K4m3_Ni_EZH2i_mean_cov", "H3K4m3_Pr_EZH2i_mean_cov",
               "H2Aub_Ni_mean_cov", "H2Aub_Pr_mean_cov",  "H2Aub_Ni_EZH2i_mean_cov", "H2Aub_Pr_EZH2i_mean_cov",
               tpm_cols)

values <- genes[genes$name %in% naive_markers, cols_used]
values[, tpm_cols] <- log2(values[, tpm_cols] + 1)

write.table(values, "./figures_data/fig2_naive_heatmap_global_scale.tsv")

fig <- combined_heatmap(
  genes,
  naive_markers,
  rnaseq_limits = c(0, 11.5),
  k4m3_limits = c(0, 120),
  k27m3_limits = c(0, 8),
  ub_limits = c(0, 11)
)

fig 

```

## Primed markers

```{r combined-summarized-heatmaps-primed, warning=F, message=F, fig.width = 7, fig.height = 8}
primed_markers <-
  c("CD24",
    "ZIC2",
    "SFRP2",
    "OTX2",
    "CYTL1",
    "HMX2",
    "THY1",
    "DUSP6",
    "PTPRZ1"
  )

values <- genes[genes$name %in% primed_markers, cols_used]
values[, tpm_cols] <- log2(values[, tpm_cols] + 1)

write.table(values, "./figures_data/fig2_primed_heatmap_global_scale.tsv")

fig <- combined_heatmap(
  genes,
  primed_markers,
  rnaseq_limits = c(0, 11.5),
  k4m3_limits = c(0, 120),
  k27m3_limits = c(0, 8),
  ub_limits = c(0, 11)
)

fig
```

# H2Aub density plot

```{r fig2d-h2aub-density-plot}
tbl <- genes %>% filter(RNASeq_DS_EZH2i_vs_Ni_log2FoldChange > 1 &
                     RNASeq_DS_EZH2i_vs_Ni_padj < 0.05) %>%
  select(H2Aub_DS_EZH2i_vs_Ni_log2FoldChange, k27_bivalency_grp) %>%
  mutate(class = ifelse(k27_bivalency_grp %in% c("Always_up", "Ni_higher_than_Pr"),
    "Bivalent_Naive", "Not"))

ggdensity(tbl, x = "H2Aub_DS_EZH2i_vs_Ni_log2FoldChange", y = "..count..",
          color="class", add = "median", fill = NA,
          palette = c("brown1","gray")) + 
  scale_x_continuous(limits=c(-2.5,2.5)) +
  labs(title = "Upregulated genes EZH2i vs Naive")
```
