---
title: "hESC data global read counts"
author: "Carmen Navarro"
date: "2021-02-03"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params:
  datadir: /home/carmen/work/publications/hesc-epigenomics/data
  fontsize: 14
---


```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(wigglescout)
library(ggplot2)
library(reshape2)
library(dplyr)
library(knitr)

knitr::opts_chunk$set(dev = c('png', 'svg'))
source("./code/embed_functions.R")
```


# Summary

This is a report on global read counts for Hu2 data.

**Pipeline version**: https://github.com/NBISweden/minute/tree/8c7646949abae12ef5bd9295f20fa6eeb7182533 .

**Mapping stats summary**: `data/meta/Kumar_2020_stats_summary.csv`.

INRC stands for Input Normalized Read Counts. Each sample number of mapped reads
is divided by the corresponding number of reads in the Input. This can be done
because of the multiplexed nature of MINUTE-ChIP protocol.

In the following plots, the value shown is the log2 enrichment over the 
reference INRC value per sample, which is the matched untreated Naïve sample
(pooled). Each dot represents one replicate.

## Helper functions

```{r helper-functions}
#' Calculate INRC from a mapped read counts table, and append such values
#' to it.
#'
#' @param counts Counts table. Corresponding file is provided as part of the 
#'   included metadata.
#' @param selector Counts column used. Final_mapped represents the final number
#'   of reads after deduplication and blacklisting.

#' @return A table including INRC and INRC norm to naive reference
calculate_inrc <- function(counts, selector = "final_mapped") {
  counts$condition <- paste(counts$celltype, counts$treatment, sep="_")
  
  inputs <- counts[counts$ip == "Input", c("library", selector)]
  colnames(inputs) <- c("library", "input_reads")
  
  non_inputs <- counts[counts$ip != "Input",]
  counts <- merge(non_inputs, inputs, by.x="input", by.y="library")
  counts$inrc <- counts[, selector] / counts[, "input_reads"]
  
  references <- counts[grepl("_Ni_pooled", counts$library), c("ip", "inrc")]
  colnames(references) <- c("ip", "ref_inrc")
  
  counts <- merge(counts, references, by="ip")
  counts$norm_to_naive <- log2(counts$inrc / counts$ref_inrc)
  
  id_vars <- c("ip", "treatment", "celltype", "condition", "replicate", "norm_to_naive")
  inrc <- counts[, c(id_vars)]
  
  inrc$condition <- factor(inrc$condition, levels = c("Naive_Untreated", "Primed_Untreated", "Naive_EZH2i", "Primed_EZH2i"))
  
  inrc
}

#' Barplot INRC pooled vs replicates per condition
#'
#' @param inrc Table with the INRC values
#' @param ip Which IP to plot
#' @param colors Corresponding colors
inrc_barplot <- function(inrc, ip, colors) {
  inrc <- inrc[inrc$ip == ip, ]
  
  max_v <- max(abs(inrc$norm_to_naive))
  
  aesthetics <- aes(x = .data[["condition"]],
                    y = .data[["norm_to_naive"]],
                    color = .data[["condition"]])
  
  ggplot(inrc[inrc$replicate!='pooled',], aesthetics) +
  geom_point() +
  geom_bar(data=inrc[inrc$replicate=='pooled',],
           stat='identity', alpha=0.6, aes(fill=condition)) +
  scale_fill_manual(values = colors) +
  scale_color_manual(values = colors) +
  labs(x = "", y = 'log2(INRC vs Naïve)', title = paste(ip, "MINUTE-ChIP")) +
  theme_classic(base_size=params$fontsize) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) + ylim(-max_v, max_v)
}

colors_list <- c("Naive_EZH2i"="#5F9EA0",
                 "Naive_Untreated"="#278b8b",
                 "Primed_EZH2i"="#f47770",
                 "Primed_Untreated"="#f44b34")

```

# Global read counts levels

## H2AUb levels

```{r h2aub-inrc-barplot}
counts_file <- file.path(params$datadir, "meta", "Kumar_2020_stats_summary.csv")
counts <- read.table(counts_file, sep="\t", header = T, na.strings = "NA", stringsAsFactors = F)

inrc <- calculate_inrc(counts) 
ip <- "H2Aub"

inrc_barplot(inrc, "H2Aub", colors_list)
```

You can download data values here: `r embed_last_plot_data(name="H2AUb_global.tsv")`.

## H3K27m3 levels

```{r h3k27m3-inrc-barplot}
inrc_barplot(inrc, "H3K27m3", colors_list)
```

You can download data values here: `r embed_last_plot_data(name="H3K27m3_global.tsv")`.

## H3K4m3 levels

```{r h3k4m3-inrc-barplot}
inrc_barplot(inrc, "H3K4m3", colors_list)
```

You can download data values here: `r embed_last_plot_data(name="H3K4m3_global.tsv")`.
