---
title: "Master bins table"
author: "Carmen Navarro"
date: "2021-05-26"
output:
  workflowr::wflow_html:
    code_folding: show
editor_options:
  chunk_output_type: console
params:
  datadir: ./data/
  pval_cutoff: 0.05
  fc_cutoff: 1
  bin_size: 10000
  shrink: FALSE
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(wigglescout)
library(ggplot2)
library(knitr)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(org.Hs.eg.db)
library(rtracklayer)
library(GenomicRanges)
library(purrr)
library(dplyr)
library(xfun)
library(cowplot)
library(tidyr)
library(DESeq2)

source("./code/ucsc_annotations.R")
source("./code/deseq_functions.R")
source("./code/embed_functions.R")
source("./code/globals.R")

knitr::opts_chunk$set(dev = c('png'))
knitr::opts_chunk$set(class.source='fold-show')

table_name <- "Kumar_2020_master_bins_10kb_table_final_raw.tsv"
table_file <- file.path("./data/meta", table_name)
table_path_html <- file.path("../data/meta", table_name)
```


This notebook shows how the master bins table is generated. Essentially, `hg38`
human genome is partitioned into `r params$bin_size` bp windows and their mean
coverage value is calculated for H3K4m3, H3K27m3 and H2AUb. DeSeq2 is applied in
a Minute-ChIP specific manner and bins are annotated as differential across
conditions:

* Primed vs Naïve,
* EZH2i treated Naïve vs Naïve and
* EZH2i treated Primed vs Primed.

Additionally, also a cross-comparison between H2Aub and H3K27m3 marks is done,
both for Naïve state and Primed state.

Final table includes these values, fold-change differences and statistical
significance scores for all genes.

* [**Master bins table**](`r table_path_html`).

# Helper functions

```{r helper-functions, class.source = NULL}

calculate_results <- function(diff, shrink, pval, coef) {
  diff_lfc <- NULL
  if (shrink == TRUE) {
    diff_lfc <- lfcShrink(diff, coef=coef, type="apeglm")
  } else {
    diff_lfc <- results(diff, alpha = pval)
  }
  diff_lfc
}

make_diff_df <- function(diff_lfc, loci, prefix) {
  df_diff <- data.frame(diff_lfc)
  # DS stands for DeSeq
  colnames(df_diff) <- paste(prefix, colnames(df_diff), sep = "_")

  cbind(data.frame(loci)[, c("seqnames", "start", "end", "strand")], df_diff)
}

make_label <- function(fnames) {
  labs <- gsub("_pooled.hg38.*scaled.bw", "", basename(fnames))
  # Remove the uncomfortable . in EZH2i elements
  labs <- gsub("-", "_", labs)
  labs <- gsub("H9_", "", labs)
  paste(labs, "mean_cov", sep = "_")
}


make_label2 <- function(fnames) {
  labs <- gsub(".hg38.*scaled.bw", "", basename(fnames))
  # Remove the uncomfortable . in EZH2i elements
  labs <- gsub("-", "_", labs)
  labs <- gsub("H9_", "", labs)
  paste(labs, "mean_cov", sep = "_")
}


gr_cbind <- function(lociset) {
  dfs <- lapply(lociset, data.frame)
  values <- dfs %>% reduce(left_join, by=c("seqnames", "start", "end", "strand", "width"))
  values
}

diff_analysis <- function(bwfiles_c1, bwfiles_c2, gr, name_c1, name_c2, shrink, pval) {
  c1 <- bw_loci(bwfiles_c1, gr)
  c2 <- bw_loci(bwfiles_c2, gr)

  diff <- bw_granges_diff_analysis(c1, c2, name_c1, name_c2,
                                   estimate_size_factors = FALSE)
  
  coef <- paste0("condition_", name_c2, "_vs_", name_c1)
  diff_lfc <- calculate_results(diff, shrink, pval, coef)
  diff_lfc
}

```

# Config analysis

```{r config-analysis}
get_bw_files <- function(pattern) {
  bwdir <- file.path(params$datadir, "bw/Kumar_2020")
  list.files(bwdir, pattern = pattern, full.names = T)
}

bins_gr <- build_bins(bin_size = params$bin_size, genome = "hg38")

# Subsample for testing
# bins_gr <- sort(bins_gr[sample(1:length(bins_gr), 2000), ])

bwfiles <-
  list(
    k4_naive = get_bw_files("H3K4m3_H9_Ni_rep[1-3].hg38.scaled.bw"),
    k4_naive_ezh2i = get_bw_files("H3K4m3_H9_Ni-EZH2i_rep[1-3].hg38.scaled.bw"),
    k4_primed = get_bw_files("H3K4m3_H9_Pr_rep[1-3].hg38.scaled.bw"),
    k4_primed_ezh2i = get_bw_files("H3K4m3_H9_Pr-EZH2i_rep[1-3].hg38.scaled.bw"),
    k27_naive = get_bw_files("H3K27m3_H9_Ni_rep[1-3].hg38.scaled.bw"),
    k27_primed = get_bw_files("H3K27m3_H9_Pr_rep[1-3].hg38.scaled.bw"),
    ub_naive = get_bw_files("H2Aub_H9_Ni_rep[1-3].hg38.scaled.bw"),
    ub_naive_ezh2i = get_bw_files("H2Aub_H9_Ni-EZH2i_rep[1-3].hg38.scaled.bw"),
    ub_primed = get_bw_files("H2Aub_H9_Pr_rep[1-3].hg38.scaled.bw"),
    ub_primed_ezh2i = get_bw_files("H2Aub_H9_Pr-EZH2i_rep[1-3].hg38.scaled.bw"),
    in_naive = get_bw_files("IN_H9_Ni.*rep[1-3].hg38.*.bw"),
    in_naive_ezh2i = get_bw_files("IN_H9_Ni-EZH2i.*rep[1-3].hg38.*.bw"),
    in_primed = get_bw_files("IN_H9_Pr_rep[1-3].hg38.*.bw"),
    in_primed_ezh2i = get_bw_files("IN_H9_Pr-EZH2i.*rep[1-3].hg38.*.bw")
  )

bwfiles_pooled <-
  list(
    k4 = get_bw_files("H3K4m3.*pooled.hg38.scaled.*"),
    k27 = get_bw_files("H3K27m3.*pooled.hg38.scaled.*"),
    ub = get_bw_files("H2Aub.*pooled.hg38.scaled.*"),
    input = get_bw_files("IN.*pooled.hg38.*")
  )

```

# Raw pooled values at bins

```{r raw-pooled, warning=F, message=F}
pooled_k4 <- bw_loci(bwfiles_pooled$k4, bins_gr, labels = make_label(bwfiles_pooled$k4))
pooled_k27 <- bw_loci(bwfiles_pooled$k27, bins_gr, labels = make_label(bwfiles_pooled$k27))
pooled_h2aub <- bw_loci(bwfiles_pooled$ub, bins_gr, labels = make_label(bwfiles_pooled$ub))
pooled_inp <- bw_loci(bwfiles_pooled$input, bins_gr, labels = make_label(bwfiles_pooled$input))

pooled_df <- gr_cbind(list(pooled_k4, pooled_k27, pooled_h2aub, pooled_inp))

master_df <- pooled_df
```

# Raw replicates values at bins

```{r raw-replicates, warning=F, message=F}
reps_df <- bw_loci(unname(unlist(bwfiles)), loci = bins_gr, labels = make_label2(unname(unlist(bwfiles))))
# In case we want the replicates as well in the master table
master_df <- gr_cbind(list(pooled_k4, pooled_k27, pooled_h2aub, pooled_inp, reps_df))
```


# K27m3 diff analysis

## Primed vs Naive

```{r k27m3-diff-pr-ni, warning=F, message=F}

diff_lfc <- diff_analysis(
  bwfiles$k27_naive,
  bwfiles$k27_primed,
  bins_gr,
  "Naive",
  "Primed",
  params$shrink,
  params$pval_cutoff
)

plotMA(diff_lfc)

df_diff <- make_diff_df(diff_lfc, bins_gr, "H3K27m3_DS_Pr_vs_Ni")

master_df <- left_join(master_df, df_diff,
                       by = c("seqnames", "start", "end", "strand"))
```

## EZH2i vs Naive and primed

These are skipped, as EZH2i treatment wipes all H3K27me3 so it does not make
any sense to do the differential analysis in this  context.


# H3K4m3 diff analysis

## Primed vs Naive

```{r k4m3-diff-pr-ni, warning=F, message=F}
diff_lfc <- diff_analysis(
  bwfiles$k4_naive,
  bwfiles$k4_primed,
  bins_gr,
  "Naive",
  "Primed",
  params$shrink,
  params$pval_cutoff
)

plotMA(diff_lfc)

df_diff <- make_diff_df(diff_lfc, bins_gr, "H3K4m3_DS_Pr_vs_Ni")

master_df <- left_join(master_df, df_diff,
                       by = c("seqnames", "start", "end", "strand"))

```


## EZH2i vs Naive

```{r k4m3-diff-ezh2i-ni, warning=F, message=F}
diff_lfc <- diff_analysis(
  bwfiles$k4_naive,
  bwfiles$k4_naive_ezh2i,
  bins_gr,
  "Naive",
  "EZH2i",
  params$shrink,
  params$pval_cutoff
)

plotMA(diff_lfc)

df_diff <- make_diff_df(diff_lfc, bins_gr, "H3K4m3_DS_EZH2i_vs_Ni")

master_df <- left_join(master_df, df_diff,
                       by = c("seqnames", "start", "end", "strand"))

```

## EZH2i vs Primed

```{r k4m3-diff-ezh2i-pr, warning=F, message=F}
diff_lfc <- diff_analysis(
  bwfiles$k4_primed,
  bwfiles$k4_primed_ezh2i,
  bins_gr,
  "Primed",
  "EZH2i",
  params$shrink,
  params$pval_cutoff
)

plotMA(diff_lfc)

df_diff <- make_diff_df(diff_lfc, bins_gr, "H3K4m3_DS_EZH2i_vs_Pr")

master_df <- left_join(master_df, df_diff,
                       by = c("seqnames", "start", "end", "strand"))

```

# H2AUb diff analysis
## Primed vs Naive

```{r h2aub-diff-pr-ni, warning=F, message=F}
diff_lfc <- diff_analysis(
  bwfiles$ub_naive,
  bwfiles$ub_primed,
  bins_gr,
  "Naive",
  "Primed",
  params$shrink,
  params$pval_cutoff
)

plotMA(diff_lfc)

df_diff <- make_diff_df(diff_lfc, bins_gr, "H2Aub_DS_Pr_vs_Ni")

master_df <- left_join(master_df, df_diff,
                       by = c("seqnames", "start", "end", "strand"))

```


## EZH2i vs Naive

```{r h2aub-diff-ezh2i-ni, warning=F, message=F}
diff_lfc <- diff_analysis(
  bwfiles$ub_naive,
  bwfiles$ub_naive_ezh2i,
  bins_gr,
  "Naive",
  "EZH2i",
  params$shrink,
  params$pval_cutoff
)

plotMA(diff_lfc)

df_diff <- make_diff_df(diff_lfc, bins_gr, "H2Aub_DS_EZH2i_vs_Ni")

master_df <- left_join(master_df, df_diff,
                       by = c("seqnames", "start", "end", "strand"))

```

## EZH2i vs Primed

```{r h2aub-diff-ezh2i-pr, warning=F, message=F}
diff_lfc <- diff_analysis(
  bwfiles$ub_primed,
  bwfiles$ub_primed_ezh2i,
  bins_gr,
  "Primed",
  "EZH2i",
  params$shrink,
  params$pval_cutoff
)

plotMA(diff_lfc)

df_diff <- make_diff_df(diff_lfc, bins_gr, "H2Aub_DS_EZH2i_vs_Pr")

master_df <- left_join(master_df, df_diff,
                       by = c("seqnames", "start", "end", "strand"))
```

# H3K27m3 vs H2AUb diff

## Naïve

```{r h2aub-diff-h3k27m3-ni, warning=F, message=F}
diff_lfc <- diff_analysis(
  bwfiles$ub_naive,
  bwfiles$k27_naive,
  bins_gr,
  "H2Aub_Ni",
  "H3K27m3_Ni",
  params$shrink,
  params$pval_cutoff
)

plotMA(diff_lfc)

df_diff <- make_diff_df(diff_lfc, bins_gr, "DS_H3K27m3_Ni_vs_H2Aub_Ni")

master_df <- left_join(master_df, df_diff,
                       by = c("seqnames", "start", "end", "strand"))
```


## Primed
```{r h2aub-diff-h3k27m3-pr, warning=F, message=F}
diff_lfc <- diff_analysis(
  bwfiles$ub_primed,
  bwfiles$k27_primed,
  bins_gr,
  "H2Aub_Pr",
  "H3K27m3_Pr",
  params$shrink,
  params$pval_cutoff
)

plotMA(diff_lfc)

df_diff <- make_diff_df(diff_lfc, bins_gr, "DS_H3K27m3_Pr_vs_H2Aub_Pr")

master_df <- left_join(master_df, df_diff,
                       by = c("seqnames", "start", "end", "strand"))
```


# Final table

```{r final-table}

write.table(
  format(master_df, digits = 4),
  file = table_file, 
  sep = "\t",
  col.names = T,
  quote = F,
  row.names = F
)

```

